---
title: "Introduction to BioGSP: Biological Graph Signal Processing"
author: "BioGSP Development Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to BioGSP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

The BioGSP package provides tools for analyzing spatial patterns in biological data using Graph Signal Processing methods, including Spectral Graph Wavelet Transform (SGWT). This vignette introduces the main concepts and demonstrates basic usage.

## Loading the Package

```{r load}
library(BioGSP)
```

## Basic Concepts

The Spectral Graph Wavelet Transform extends classical wavelet analysis to irregular domains by:

1. **Graph Construction**: Building a graph from spatial coordinates
2. **Spectral Analysis**: Using graph Laplacian eigendecomposition
3. **Wavelet Design**: Creating wavelets in the spectral domain
4. **Multi-scale Decomposition**: Analyzing signals at multiple scales

## Simple Example

Let's start with a simple synthetic dataset:

```{r synthetic_data}
# Generate synthetic spatial data
set.seed(123)
n_points <- 64
x_coords <- rep(1:8, each = 8) + rnorm(n_points, 0, 0.1)
y_coords <- rep(1:8, times = 8) + rnorm(n_points, 0, 0.1)

# Create a spatial pattern
signal_data <- sin(0.8 * x_coords) * cos(0.8 * y_coords) + 
               0.3 * sin(1.5 * x_coords * y_coords) + 
               rnorm(n_points, 0, 0.1)

demo_data <- data.frame(x = x_coords, y = y_coords, signal = signal_data)

# View the data structure
head(demo_data)
```

## SGWT Analysis

Now let's apply SGWT to analyze this spatial signal:

```{r sgwt_analysis}
# Apply SGWT
result <- SGWT(data.in = demo_data, 
               signal = "signal",
               k = 4,               # number of nearest neighbors
               J = 3,               # number of scales
               scaling_factor = 2,
               k_fold = 5)          # k_fold * sqrt(64) = 5 * 8 = 40 < 64

# View reconstruction quality
cat("Reconstruction RMSE:", result$reconstruction_error, "\n")
```

## Energy Analysis

We can analyze how energy is distributed across different scales:

```{r energy_analysis}
# Analyze energy distribution
energy_df <- sgwt_energy_analysis(result)
print(energy_df)

# Plot energy distribution
library(ggplot2)
ggplot(energy_df, aes(x = scale, y = energy_ratio)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Energy Distribution Across Scales",
       x = "Scale", y = "Energy Ratio") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Visualization

The package provides visualization functions to explore the decomposition:

```{r visualization, eval=FALSE}
# Visualize SGWT decomposition (requires ggpubr)
if (requireNamespace("ggpubr", quietly = TRUE)) {
  plots <- plot_sgwt_decomposition(result, demo_data, plot_scales = 1:3)
  print(plots)
}
```

## Advanced Usage

### Custom Scales

You can specify custom scales for analysis:

```{r custom_scales}
# Define custom scales
custom_scales <- c(1.5, 0.8, 0.4, 0.2)

result_custom <- SGWT(data.in = demo_data,
                      signal = "signal",
                      scales = custom_scales,
                      k = 4,
                      k_fold = 5)

cat("Custom scales reconstruction RMSE:", result_custom$reconstruction_error, "\n")
```

### Different Laplacian Types

You can use different types of graph Laplacians:

```{r laplacian_types}
# Normalized Laplacian (default)
result_norm <- SGWT(demo_data, "signal", k = 4, J = 3, laplacian_type = "normalized", k_fold = 5)

# Unnormalized Laplacian
result_unnorm <- SGWT(demo_data, "signal", k = 4, J = 3, laplacian_type = "unnormalized", k_fold = 5)

cat("Normalized Laplacian RMSE:", result_norm$reconstruction_error, "\n")
cat("Unnormalized Laplacian RMSE:", result_unnorm$reconstruction_error, "\n")
```

## Graph Cross-Correlation

The package also provides tools for analyzing relationships between spatial signals:

```{r gcc_example}
# Create a second signal
demo_data$signal2 <- cos(0.6 * x_coords) * sin(0.6 * y_coords) + rnorm(n_points, 0, 0.1)

# Calculate eigendecomposition
eigen_result <- Cal_Eigen(demo_data, k = 4, k_fold = 5, sensitivity = 2)

# Calculate cross-correlation
gcc_value <- Cal_GCC(data.in = demo_data,
                     knee = eigen_result[[1]],
                     signal1 = "signal",
                     signal2 = "signal2",
                     eigenvector = eigen_result[[2]])

cat("Graph Cross-Correlation:", gcc_value, "\n")
```

## Demo Function

The package includes a demo function for quick testing:

```{r demo}
# Run the built-in demo
demo_result <- demo_sgwt()

# View demo results
print(demo_result$energy)
```

## Conclusion

The SGWT package provides a comprehensive toolkit for multi-scale analysis of spatial data. Key features include:

- Flexible graph construction from spatial coordinates
- Multiple wavelet kernel options
- Energy analysis across scales
- Visualization tools
- Cross-correlation analysis

This makes it particularly useful for applications in spatial biology, image processing, and network analysis.

## References

Hammond, D. K., Vandergheynst, P., & Gribonval, R. (2011). Wavelets on graphs via spectral graph theory. Applied and Computational Harmonic Analysis, 30(2), 129-150. 