---
title: "BioGSP: Spectral Graph Wavelet Transform for Spatial Data (New Workflow)"
author: "Yuzhou"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
vignette: >
  %\VignetteIndexEntry{BioGSP: Spectral Graph Wavelet Transform for Spatial Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      fig.width = 10, fig.height = 6)
```

# Introduction

BioGSP provides **Spectral Graph Wavelet Transform (SGWT)** analysis for spatial biological data. This package enables multi-scale analysis of spatial patterns using graph signal processing techniques.


## Key Capabilities

- **Object-oriented SGWT workflow** with clear separation of concerns
- **Multi-scale spatial analysis** using spectral graph wavelets
- **Flexible data input** with custom column naming
- **Energy-normalized similarity analysis** with comprehensive metrics
- **Multiple wavelet kernels** (Mexican Hat, Meyer, Heat)
- **Comprehensive visualization** tools
- **Detailed inverse transform results** (low-pass, band-pass approximations)

## New Workflow Overview

The new workflow consists of five main steps:

1. **`initSGWT()`**: Initialize SGWT object with data and parameters
2. **`checkKband()`**: Analyze k-band limited property of signals (optional)
3. **`runSpecGraph()`**: Build spectral graph structure
4. **`runSGWT()`**: Perform forward and inverse SGWT transforms
5. **`runSGCC()`**: Calculate energy-normalized weighted similarity

# Setup

```{r load-libraries}
# Load required packages
library(ggplot2)
library(patchwork)
library(viridis)
library(Matrix)
library(igraph)
library(RANN)
library(RSpectra)

# Load BioGSP functions
source('../R/simulation.R')
source('../R/sgwt_core.R')
source('../R/sgwt_main.R')
source('../R/utilities.R')
source('../R/visualization.R')

# Set random seed for reproducibility
set.seed(123)
```

# New SGWT Workflow Demonstration

## Generate Example Pattern

```{r generate-pattern}
# Create a spatial pattern with concentric circles
demo_pattern <- simulate_multiscale(
  grid_size = 40,        # Moderate size for fast computation
  n_centers = 1,         # Single center pattern
  Ra_seq = 5,            # Inner circle radius
  n_steps = 1,           # Single step (one pattern)
  outer_start = 10,      # Outer ring radius
  seed = 123
)[[1]]

# Display pattern structure
cat("Generated pattern dimensions:", nrow(demo_pattern), "x", ncol(demo_pattern), "\n")
cat("Column names:", paste(colnames(demo_pattern), collapse = ", "), "\n")
cat("Signals available:", sum(demo_pattern$signal_1), "inner pixels,", sum(demo_pattern$signal_2), "outer pixels\n")
```

## K-band Limited Analysis

```{r kband-analysis}
# Demonstrate k-band limited property analysis
cat("=== K-band Limited Property Analysis ===\n")

# Initialize SGWT object for k-band analysis
SG_kband <- initSGWT(
  data.in = demo_pattern,
  x_col = "X",
  y_col = "Y", 
  signals = c("signal_1", "signal_2")
)

# Run k-band analysis (builds graph internally)
kband_result <- checkKband(
  SG = SG_kband,
  signals = c("signal_1", "signal_2"),
  k = 15,                    # Number of neighbors for graph construction
  alpha = 0.05,              # Significance level for Wilcoxon test
  verbose = FALSE
)

# Display results
cat("\n=== K-band Analysis Results ===\n")
cat("All signals are k-band limited:", kband_result$is_kband_limited, "\n")
cat("Low-frequency knee point:", kband_result$knee_point_low, "\n")
```

## Visualize Input Pattern

```{r visualize-pattern, fig.width=10, fig.height=4}
# Create visualizations for both signals
p1 <- ggplot(demo_pattern, aes(X, Y, fill = factor(signal_1))) +
  geom_tile() +
  scale_fill_manual(values = c("0" = "white", "1" = "#e31a1c"), 
                    name = "Signal", labels = c("Background", "Inner Circle")) +
  coord_fixed() + theme_void() + ggtitle("Signal 1: Inner Circles")

p2 <- ggplot(demo_pattern, aes(X, Y, fill = factor(signal_2))) +
  geom_tile() +
  scale_fill_manual(values = c("0" = "white", "1" = "#1f78b4"), 
                    name = "Signal", labels = c("Background", "Outer Ring")) +
  coord_fixed() + theme_void() + ggtitle("Signal 2: Outer Rings")

print(p1 + p2)
```

## Step 1: Initialize SGWT Object

```{r init-sgwt}
# Initialize SGWT object with custom column names
SG <- initSGWT(
  data.in = demo_pattern,
  x_col = "X",           # Custom X coordinate column name
  y_col = "Y",           # Custom Y coordinate column name
  signals = c("signal_1", "signal_2"),  # Analyze both signals
  J = 2,                 # Number of wavelet scales
  scaling_factor = 1,    # Scale progression factor
  kernel_type = "mexican_hat"
)

# Display initialized object
print(SG)
```

## Step 2: Build Spectral Graph

```{r build-graph}
# Build spectral graph structure
SG <- runSpecGraph(SG, k = 12, laplacian_type = "normalized", length_eigenvalue = 200, verbose = TRUE)

# Check updated object
cat("Graph construction completed!\n")
cat("Adjacency matrix dimensions:", dim(SG$Graph$adjacency_matrix), "\n")
cat("Number of eigenvalues computed:", length(SG$Graph$eigenvalues), "\n")
```

## Visualize Fourier Modes (Eigenvectors)

```{r fourier-modes, fig.width=12, fig.height=8}
# Visualize Fourier modes (eigenvectors) - 5 low and 5 high frequency modes
fourier_modes <- plot_FM(SG, mode_type = "both", n_modes = 5, ncol = 5)

cat("Fourier Modes Visualization:\n")
cat("- Low-frequency modes (2-6): Smooth spatial patterns, excluding DC component\n")
cat("- High-frequency modes (96-100): Fine-detailed spatial patterns\n")
cat("- Each mode shows its eigenvalue (λ) indicating frequency content\n")
```

## Step 3: Run SGWT Analysis

```{r run-sgwt}
# Perform SGWT forward and inverse transforms
SG <- runSGWT(SG, verbose = TRUE)

# Display final object with all results
print(SG)
```

## Visualize SGWT Decomposition

```{r plot-decomposition, fig.width=12, fig.height=8}
# Plot SGWT decomposition results for signal_1
decomp_plots <- plot_sgwt_decomposition(
  SG = SG,
  signal_name = "signal_1",
  plot_scales = 1:3,     # Show first 3 scales
  ncol = 3
)

```

## Energy Analysis

```{r energy-analysis}
# Analyze energy distribution across scales for signal_1
energy_analysis <- sgwt_energy_analysis(SG, "signal_1")
print(energy_analysis)

# Create energy distribution plot
energy_plot <- ggplot(energy_analysis, aes(x = scale, y = energy_ratio)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.7) +
  geom_text(aes(label = paste0(round(energy_ratio * 100, 1), "%")), 
            vjust = -0.5, size = 3) +
  labs(title = "Energy Distribution Across SGWT Scales (Signal 1)",
       x = "Scale", y = "Energy Ratio") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(energy_plot)
```

# Advanced Features Demonstration

## Different Wavelet Kernels

```{r kernel-comparison}
# Compare different kernel types
kernels <- c("mexican_hat", "meyer", "heat")
kernel_results <- list()

for (kn in kernels) {
  # Initialize with different kernel
  SG_temp <- initSGWT(
    data.in = demo_pattern,
    x_col = "X", y_col = "Y",
    signals = "signal_1",
    J = 3,                    # Reduced for faster computation
    kernel_type = kn
  )
  
  # Run full workflow
  SG_temp <- runSpecGraph(SG_temp, k = 12, laplacian_type = "normalized", length_eigenvalue = 100,verbose = FALSE)
  SG_temp <- runSGWT(SG_temp, verbose = FALSE)
  
  kernel_results[[kn]] <- SG_temp
}

# Compare reconstruction errors
comparison_df <- data.frame(
  Kernel = kernels,
  RMSE = sapply(kernel_results, function(x) x$Inverse$signal_1$reconstruction_error),
  stringsAsFactors = FALSE
)

print("Kernel Performance Comparison:")
print(comparison_df)

# Plot comparison
ggplot(comparison_df, aes(x = Kernel, y = RMSE, fill = Kernel)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_text(aes(label = round(RMSE, 6)), vjust = -0.5) +
  scale_fill_viridis_d() +
  labs(title = "SGWT Reconstruction Error by Kernel Type",
       x = "Kernel Type", y = "RMSE") +
  theme_minimal()
```

## Step 4: Pattern Similarity Analysis

```{r similarity-analysis}
# Calculate similarity between signal_1 and signal_2 in same object
similarity_within <- runSGCC("signal_1", "signal_2", SG = SG, return_parts = TRUE)

cat("Pattern Similarity Analysis (within same graph):\n")
cat(sprintf("Overall similarity: %.4f\n", similarity_within$S))
cat(sprintf("Low-frequency similarity: %.4f\n", similarity_within$c_low))
cat(sprintf("Non-low-frequency similarity: %.4f\n", similarity_within$c_nonlow))
cat(sprintf("Energy weights - Low: %.3f, Non-low: %.3f\n", 
           similarity_within$w_low, similarity_within$w_NL))

# Generate a second pattern for cross-comparison
pattern_2 <- simulate_multiscale(
  grid_size = 40,
  n_centers = 1,
  Ra_seq = 8,            # Different inner radius
  n_steps = 1,           # Single step
  outer_start = 15,      # Different outer radius
  seed = 456
)[[1]]

# Create second SGWT object
SG2 <- initSGWT(pattern_2, x_col = "X", y_col = "Y", signals = "signal_1", 
                J = 4)
SG2 <- runSpecGraph(SG2, k = 12, laplacian_type = "normalized",length_eigenvalue = 100, verbose = FALSE)
SG2 <- runSGWT(SG2, verbose = FALSE)

# Note: Cross-object similarity comparison removed due to different eigenvalue counts
# between SG (200 eigenvalues) and SG2 (100 eigenvalues) causing dimension mismatch

# Visualize both patterns for comparison
p_comp1 <- ggplot(demo_pattern, aes(X, Y, fill = factor(signal_1))) +
  geom_tile() + scale_fill_manual(values = c("0" = "white", "1" = "#e31a1c")) +
  coord_fixed() + theme_void() + theme(legend.position = "none") +

    ggtitle("Pattern A (R_in=5, R_out=10)")

p_comp2 <- ggplot(pattern_2, aes(X, Y, fill = factor(signal_1))) +
  geom_tile() + scale_fill_manual(values = c("0" = "white", "1" = "#1f78b4")) +
  coord_fixed() + theme_void() + theme(legend.position = "none") +
  ggtitle("Pattern B (R_in=7, R_out=12)")

print(p_comp1 + p_comp2)
```

## Low-frequency Only Analysis

```{r low-frequency-analysis}
# Compare using only low-frequency components
similarity_low <- runSGCC("signal_1", "signal_2", SG = SG, low_only = TRUE, return_parts = TRUE)

cat("Low-frequency Only Similarity Analysis:\n")
cat(sprintf("Low-frequency similarity: %.4f\n", similarity_low$c_low))
cat(sprintf("Overall score (same as low-freq): %.4f\n", similarity_low$S))
cat("Note: Non-low-frequency components are ignored (c_nonlow = NA)\n")

# Compare with full analysis
cat("\nComparison:\n")
cat(sprintf("Full analysis similarity: %.4f\n", similarity_within$S))
cat(sprintf("Low-only similarity: %.4f\n", similarity_low$S))
cat(sprintf("Difference: %.4f\n", abs(similarity_within$S - similarity_low$S)))
```

## Multiple Signal Analysis

```{r multiple-signals}
# Analyze energy distribution for both signals
energy_signal1 <- sgwt_energy_analysis(SG, "signal_1")
energy_signal2 <- sgwt_energy_analysis(SG, "signal_2")

# Combine for comparison
energy_comparison <- rbind(energy_signal1, energy_signal2)

# Plot comparison
ggplot(energy_comparison, aes(x = scale, y = energy_ratio, fill = signal)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  geom_text(aes(label = paste0(round(energy_ratio * 100, 1), "%")), 
            position = position_dodge(width = 0.9), vjust = -0.5, size = 3) +
  scale_fill_viridis_d() +
  labs(title = "Energy Distribution Comparison: Signal 1 vs Signal 2",
       x = "Scale", y = "Energy Ratio", fill = "Signal") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Similarity Space Visualization with Multiple Patterns

```{r similarity-visualization, fig.width=12, fig.height=10}
# Generate 9 paired patterns using simulate_multiscale
cat("Generating 9 paired patterns for similarity analysis...\n")

patterns_9 <- simulate_multiscale(
  grid_size = 40,
  n_centers = 1,
  Ra_seq = c(3, 6, 9),     # Inner circle radii
  n_steps = 3,             # 3 shrinkage steps
  outer_start = 20,        # Fixed starting outer radius
  seed = 123
)

cat("Generated", length(patterns_9), "patterns\n")
cat("Pattern names:", names(patterns_9), "\n")

# Visualize all 9 patterns first
pattern_viz <- visualize_multiscale(
  patterns_9, 
  Ra_seq = c(3, 6, 9), 
  n_steps = 3,
  show_subtitle = TRUE
)
print(pattern_viz)

# Create SGWT objects for all 9 patterns and compute similarities
similarity_results <- list()
sgwt_objects <- list()

cat("\nProcessing patterns and computing SGWT analysis...\n")

# Process each pattern
for (i in seq_along(patterns_9)) {
  pattern_name <- names(patterns_9)[i]
  pattern_data <- patterns_9[[i]]
  
  cat("Processing", pattern_name, "...\n")
  
  # Create SGWT object
  SG_temp <- initSGWT(
    data.in = pattern_data,
    x_col = "X", 
    y_col = "Y", 
    signals = c("signal_1", "signal_2"),
    J = 4,
    kernel_type = "heat"
  )
  
  # Build graph and run SGWT
  SG_temp <- runSpecGraph(SG_temp, k = 12, laplacian_type = "normalized", 
                         length_eigenvalue = 50, verbose = FALSE)
  SG_temp <- runSGWT(SG_temp, verbose = FALSE)
  
  sgwt_objects[[pattern_name]] <- SG_temp
  
  # Compute within-pattern similarity (signal_1 vs signal_2)
  sim_within <- runSGCC("signal_1", "signal_2", SG = SG_temp, return_parts = TRUE)
  similarity_results[[pattern_name]] <- sim_within
}

# Print similarity results summary
cat("\nSimilarity Results Summary (signal_1 vs signal_2 within each pattern):\n")
similarity_df <- data.frame(
  Pattern = names(similarity_results),
  S = sapply(similarity_results, function(x) x$S),
  c_low = sapply(similarity_results, function(x) x$c_low),
  c_nonlow = sapply(similarity_results, function(x) x$c_nonlow),
  w_low = sapply(similarity_results, function(x) x$w_low),
  w_NL = sapply(similarity_results, function(x) x$w_NL),
  stringsAsFactors = FALSE
)

print(similarity_df)

# Extract Ra and Step values for better labeling
similarity_df$Ra <- as.numeric(gsub(".*Ra_([0-9.]+)_Step.*", "\\1", similarity_df$Pattern))
similarity_df$Step <- as.numeric(gsub(".*Step_([0-9]+)$", "\\1", similarity_df$Pattern))
similarity_df$Label <- paste0("Ra=", similarity_df$Ra, ",Step=", similarity_df$Step)

# Create similarity space visualization
similarity_plot <- visualize_similarity_xy(
  similarity_results,
  point_size = 4,
  point_color = "steelblue",
  add_diagonal = TRUE,
  add_axes_lines = TRUE,
  title = "SGWT Similarity Space: 9 Paired Patterns (signal_1 vs signal_2)",
  show_labels = TRUE
)

print(similarity_plot)

# Create enhanced plot with color coding by Ra values
library(ggplot2)
plot_data <- data.frame(
  c_low = similarity_df$c_low,
  c_nonlow = similarity_df$c_nonlow,
  Ra = factor(similarity_df$Ra),
  Step = factor(similarity_df$Step),
  Label = similarity_df$Label,
  S = similarity_df$S
)

enhanced_plot <- ggplot(plot_data, aes(x = c_low, y = c_nonlow, color = Ra, shape = Step)) +
  geom_point(size = 5, alpha = 0.8) +
  geom_text(aes(label = Label), vjust = -0.8, hjust = 0.5, size = 3, color = "black") +
  xlim(-1, 1) + ylim(-1, 1) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray60", alpha = 0.7) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray60", alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", color = "gray40", alpha = 0.7) +
  geom_abline(slope = -1, intercept = 0, linetype = "dotted", color = "gray40", alpha = 0.7) +
  scale_color_viridis_d(name = "Inner Radius (Ra)") +
  scale_shape_manual(values = c(16, 17, 15), name = "Shrinkage Step") +
  labs(
    title = "Enhanced Similarity Space: Color by Ra, Shape by Step",
    x = "Low-frequency Similarity (c_low)",
    y = "Non-low-frequency Similarity (c_nonlow)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.position = "right"
  )

print(enhanced_plot)

# Analysis of patterns
cat("\nPattern Analysis:\n")
cat("Ra (Inner Radius) Effect:\n")
for (ra in unique(similarity_df$Ra)) {
  subset_data <- similarity_df[similarity_df$Ra == ra, ]
  cat(sprintf("  Ra=%g: Mean c_low=%.3f, Mean c_nonlow=%.3f, Mean S=%.3f\n", 
              ra, mean(subset_data$c_low), mean(subset_data$c_nonlow), mean(subset_data$S)))
}

cat("\nStep (Shrinkage Step) Effect:\n")
for (step in unique(similarity_df$Step)) {
  subset_data <- similarity_df[similarity_df$Step == step, ]
  cat(sprintf("  Step=%g: Mean c_low=%.3f, Mean c_nonlow=%.3f, Mean S=%.3f\n", 
              step, mean(subset_data$c_low), mean(subset_data$c_nonlow), mean(subset_data$S)))
}

cat("\nSimilarity Space Interpretation:\n")
cat("- Each point represents similarity between signal_1 (inner circle) and signal_2 (outer ring)\n")
cat("- Color indicates inner radius (Ra): different colors for different inner radii\n")
cat("- Shape indicates shrinkage step: circle/triangle/square for different steps\n")
cat("- Position shows frequency-domain similarity characteristics\n")
```

## Moving Circles Pattern Analysis

```{r moving-circles-demo, fig.width=12, fig.height=8}
# Generate moving circles pattern
cat("Generating moving circles pattern...\n")
moving_circles_data <- simulate_moving_circles(
  grid_size = 40,
  radius_seq = c(6, 10, 14),  # Three different radii
  n_steps = 5,                # Five movement steps
  center_distance = 20,       # Initial distance from center
  radius2_factor = 1.3,       # Second circle is 1.3x larger
  seed = 789
)

cat("Moving circles dimensions:", length(moving_circles_data), "patterns\n")
cat("Pattern names:", head(names(moving_circles_data), 3), "...\n")

# Visualize the moving circles pattern
moving_viz <- visualize_moving_circles(
  moving_circles_data,
  show_subtitle = TRUE
)
print(moving_viz + ggtitle("Moving Circles Pattern: Two Circles Approaching Each Other"))

# Run SGWT analysis on one moving circles pattern
cat("\nRunning SGWT analysis on moving circles pattern...\n")

# Select one pattern for detailed analysis
selected_pattern <- moving_circles_data[["Ra_10_Step_3"]]  # Middle radius, middle step

# Initialize SGWT object
SG_moving <- initSGWT(
  data.in = selected_pattern,
  x_col = "X", 
  y_col = "Y", 
  signals = c("signal_1", "signal_2"),
  J = 4,
  kernel_type = "heat"
)

# Build spectral graph
SG_moving <- runSpecGraph(SG_moving, k = 12, laplacian_type = "normalized", 
                         length_eigenvalue = 50, verbose = FALSE)

# Run SGWT transforms
SG_moving <- runSGWT(SG_moving, verbose = FALSE)

# Calculate similarity between the two moving circles
similarity_moving <- runSGCC("signal_1", "signal_2", SG = SG_moving, return_parts = TRUE)

cat("\nMoving Circles Pattern Similarity Analysis:\n")
cat("==========================================\n")
cat(sprintf("Weighted similarity (S): %.6f\n", similarity_moving$S))
cat(sprintf("Low-frequency similarity (c_low): %.6f\n", similarity_moving$c_low))
cat(sprintf("Non-low-frequency similarity (c_nonlow): %.6f\n", similarity_moving$c_nonlow))
cat(sprintf("Energy weights - Low: %.3f, Non-low: %.3f\n", 
           similarity_moving$w_low, similarity_moving$w_NL))

cat("\nInterpretation:\n")
cat("- Moving circles create dynamic spatial patterns with mutual exclusion\n")
cat("- Signal assignment priority ensures no overlap (signal_1 takes precedence)\n")
cat("- Similarity reflects spatial relationship between approaching circles\n")

# Visualize the selected pattern
p_selected <- ggplot(selected_pattern, aes(X, Y, fill = factor(signal_1 + signal_2 * 2))) +
  geom_tile() + 
  scale_fill_manual(values = c("0" = "white", "1" = "#16964a", "2" = "#2958a8"), 
                    name = "Signal", labels = c("Background", "Circle 1", "Circle 2")) +
  coord_fixed() + theme_void() + 
  ggtitle("Selected Moving Circles Pattern (Ra=10, Step=3)") +
  theme(legend.position = "bottom")

print(p_selected)
```

# Advanced Workflow Features

## Custom Parameters

```{r custom-parameters}
# Demonstrate advanced parameter customization
SG_advanced <- initSGWT(
  data.in = demo_pattern,
  x_col = "X", y_col = "Y",
  signals = "signal_1",
  J = 6,                     # More scales for finer analysis
  scaling_factor = 1.5,      # Closer scales
  kernel_type = "heat"       # Heat kernel
)

SG_advanced <- runSpecGraph(SG_advanced, k = 15, laplacian_type = "randomwalk",length_eigenvalue = 30, verbose = FALSE)
SG_advanced <- runSGWT(SG_advanced, verbose = FALSE)

cat("Advanced Parameters Results:\n")
cat("Number of scales:", length(SG_advanced$Parameters$scales), "\n")
cat("Scales:", paste(round(SG_advanced$Parameters$scales, 4), collapse = ", "), "\n")
cat("Reconstruction error:", round(SG_advanced$Inverse$signal_1$reconstruction_error, 6), "\n")
```

## Checkerboard Pattern Analysis

```{r checkerboard-demo, fig.width=12, fig.height=8}
# Generate checkerboard pattern
cat("Generating checkerboard pattern...\n")
checkerboard_data <- simulate_checkerboard(grid_size = 10, tile_size = 1)

cat("Checkerboard dimensions:", nrow(checkerboard_data), "points\n")
cat("Signal 1 tiles:", sum(checkerboard_data$signal_1), "\n")
cat("Signal 2 tiles:", sum(checkerboard_data$signal_2), "\n")

# Visualize the checkerboard pattern
p_checker <- visualize_checkerboard(checkerboard_data, color1 = "black", color2 = "lightgray")
print(p_checker + ggtitle("Checkerboard Pattern: Alternating Signals"))

# Run SGWT analysis on checkerboard
cat("\nRunning SGWT analysis on checkerboard pattern...\n")

# Initialize SGWT object
SG_checker <- initSGWT(
  data.in = checkerboard_data,
  x_col = "X", 
  y_col = "Y", 
  signals = c("signal_1", "signal_2"),
  J = 4,
  kernel_type = "heat"
)

# Build spectral graph
SG_checker <- runSpecGraph(SG_checker, k = 12, laplacian_type = "normalized", 
                          length_eigenvalue = 50, verbose = FALSE)

# Run SGWT transforms
SG_checker <- runSGWT(SG_checker, verbose = FALSE)

# Calculate similarity between the two checkerboard signals
similarity_checker <- runSGCC("signal_1", "signal_2", SG = SG_checker, return_parts = TRUE)

cat("\nCheckerboard Pattern Similarity Analysis:\n")
cat("=====================================\n")
cat(sprintf("Weighted similarity (S): %.6f\n", similarity_checker$S))
cat(sprintf("Low-frequency similarity (c_low): %.6f\n", similarity_checker$c_low))
cat(sprintf("Non-low-frequency similarity (c_nonlow): %.6f\n", similarity_checker$c_nonlow))
cat(sprintf("Energy weights - Low: %.3f, Non-low: %.3f\n", 
           similarity_checker$w_low, similarity_checker$w_NL))

cat("\nInterpretation:\n")
cat("- Checkerboard patterns are spatially complementary (non-overlapping)\n")
cat("- High-frequency content dominates due to sharp tile boundaries\n")
cat("- Low similarity values expected due to alternating pattern structure\n")
```

# Summary

## Key Takeaways

1. **New Object-Oriented Workflow**: Clear separation of initialization, graph building, analysis, and similarity computation
2. **Flexible Input**: Supports custom column naming for coordinates and signals  
3. **Multiple Kernels**: Choose from Mexican Hat, Meyer, or Heat kernel families
4. **Comprehensive Similarity**: Energy-normalized weighted similarity with detailed diagnostics
5. **Advanced Analysis**: Low-frequency only analysis, cross-object comparison, multiple signal support

## New Workflow Benefits

```r
# Complete workflow in 5 clear steps:
SG <- initSGWT(data, signals = c("signal1", "signal2"))           # 1. Initialize
kband <- checkKband(SG, signals = c("signal1", "signal2"))       # 2. Check k-band (optional)
SG <- runSpecGraph(SG, k = 12, laplacian_type = "normalized")    # 3. Build graph  
SG <- runSGWT(SG)                                                 # 4. Run SGWT
similarity <- runSGCC("signal1", "signal2", SG = SG)             # 5. Compare patterns
```

## Parameter Guidelines

- **Small datasets (<200 points)**: k=8-12, J=3-4
- **Medium datasets (200-1000 points)**: k=12-20, J=4-6  
- **Large datasets (>1000 points)**: k=15-25, J=5-8

## Next Steps

- Explore your own spatial datasets using the demonstrated workflow
- Experiment with different kernel types and parameters
- Use comprehensive similarity analysis for comparative studies
- Refer to function documentation for detailed parameter specifications

---

*BioGSP Package - Biological Graph Signal Processing for Spatial Data Analysis*
